---
title: "Analysis_2"
author: "Manqing Lin"
date: "2024-03-27"
output: html_document
---

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

```

```{r}
df_flavor <- read_csv("/Users/linmatch/Desktop/RA/data/df_flavor.csv")
#View(df_flavor)
```

```{r}
data_long <- df_flavor %>%
  select(-c(Blinding_Code, Sample_Position)) %>%
  pivot_longer(cols = starts_with("time_"),
               names_to = "Time",
               names_prefix = "time_", 
               values_to = "Value") %>%
  mutate(Time = as.numeric(str_remove(Time, "s")),
         Sample_ID = as.factor(Sample_ID),
         Sample_Name = ifelse(Sample_Name == "Starch-Limonene Inclusion Complex",
                                "Starch-Limonene IC",
                                Sample_Name),
         Sample_Name = as.factor(Sample_Name),
         Attribute = as.factor(Attribute))
```

```{r}
data_summary <- data_long %>%
  group_by(Sample_Name, Attribute) %>%
  summarise(mean = mean(Value, na.rm = TRUE),
            sd = sd(Value, na.rm = TRUE),
            .groups = "drop")  # drop grouping afterwards
```

```{r}
data_summary <- data_long %>%
  group_by(Sample_Name, Attribute, Time) %>%
  summarise(mean_value = mean(Value, na.rm = TRUE), .groups = "drop")
```

```{r}
data_sweet_30s <- data_long %>%
  filter(Time == 30, Attribute == "Sweet")

# Extract the data for the two samples
data_free_limonene <- data_sweet_30s %>% filter(Sample_Name == "Free Limonene")
data_starch_limonene_ic <- data_sweet_30s %>% filter(Sample_Name == "Starch-Limonene IC")

# Perform the two-proportion z-test
z_test <- prop.test(x = c(sum(data_free_limonene$Value), sum(data_starch_limonene_ic$Value)),
                    n = c(nrow(data_free_limonene), nrow(data_starch_limonene_ic)),
                    alternative = "two.sided")

# Print the results
print(z_test)
```


# What have been done:
#### Visualization: 
1. Correlation Matrix
2. Comparison of Flavor Perception Between Gum Types(Side-by-Side plot)
3. Difference of Flavor Perception Between Gum Types
4. Comparison of Flavor Perception Between Gum Types by different copy number group
5. Comparison of type of change in Flavor Perception by different group("CNV_group", "sAA_After_Group" etc.)

#### Dim Reduction:
1. PCA

# More to do:
## Statistical test:
1. proportion test: Test if there is significant difference in proportion of certain flavor perception for 2 sample gums at certain time point of interest (e.g., time_1s, time_2s, etc.). 
2. proportion test: Test if there is significant difference in proportion of certain flavor perception for different gene copy numbers group (low vs. high) with certain sample gums (Free Limonene or Starch-Limonene Inclusion Complex) at certain time point of interest (e.g., time_1s, time_2s, etc.).
3. Chi-square test of independence: Test if flavor perception (yes/no) and gum type are independent or if there is an association between them.

## Visualization:
1. How sample position affect the flavor perception trend?

## Statistical Modeling:
1. Apply a generalized linear mixed-effects model (GLMM) or a generalized estimating equation (GEE) model to explore the relationship between certain flavor perception at certain time points and predictors such as gene copy number, gum type, time and their interaction term etc. 
-GLMM: log(p_ijk / (1 - p_ijk)) = β_0 + β_1 * gene_copy_num_i + β_2 * gum_type_j + β_3 * time_k
+ β_4 * (gene_copy_num_i × gum_type_j)
+ β_5 * (gene_copy_num_i × time_k)
+ β_6 * (gum_type_j × time_k)
+ β_7 * (gene_copy_num_i × gum_type_j × time_k)
+ u_i + v_ij

where:
p_ijk is the probability of perceiving the flavor for participant i with gene copy number j and gum type k at time t
β_0 is the overall intercept
β_1, β_2, β_3 are the main effects of gene copy number, gum type, and time
β_4, β_5, β_6, β_7 are the two-way and three-way interaction effects
u_i is the random intercept for participant i
v_ij is the random slope for the interaction between participant i and gum type j
