---
title: "Analysis_2"
author: "Manqing Lin"
date: "2024-03-27"
output: html_document
---

```{r}
library(readr)
```

```{r}
df_flavor <- read_csv("data/df_flavor.csv")
#View(df_flavor)
```

# What have been done:
#### Visualization: 
1. Correlation Matrix
2. Comparison of Flavor Perception Between Gum Types(Side-by-Side plot)
3. Difference of Flavor Perception Between Gum Types
4. Comparison of Flavor Perception Between Gum Types by different copy number group
5. Comparison of type of change in Flavor Perception by different group("CNV_group", "sAA_After_Group" etc.)

#### Dim Reduction:
1. PCA

# More to do:
## Statistical test:
1. t-test: Test if there is significant difference of certain flavor perception for 2 sample gums at certain time point of interest (e.g., time_1s, time_2s, etc.). 
2. t-test: Test if there is significant difference of certain flavor perception for different gene copy numbers group (low vs. high) with certain sample gums (Free Limonene or Starch-Limonene Inclusion Complex) at certain time point of interest (e.g., time_1s, time_2s, etc.).

## Visualization:
1. How sample position affect the flavor perception trend

## Prediction & Inference:
1. Use time series forecasting models (e.g., ARIMA) to predict future flavor perception trends based on historical data.
2. Apply regression model to explore the relationship between certain flavor perception at certain time points and predictors such as gene copy number, age, gender, race etc. 
